### 2023-10-09 doc/_bayes_priors.qmd
change from BayesFitness to BarBay

With this in mind, we leave it to the reader to judge the selection of priors.
Furthermore, the software package associated with this work, `BayesFitness.jl`,
is written so that users can experiment with different prior selection criteria
that fit their needs. We strongly advocate that statistics should not be done in
a black-box fit-all tool mindset but rather as a formal way to encode the
assumptions behind the analysis, subject to constructive criticism. With this
philosophical baggage behind us, let us now focus on how the priors used for
this work were selected.

@@

The `BayesFitness.jl` package includes a function `naive_prior` within the
`stats` module. This function utilizes the data from the neutral lineages to
determine some of the prior parameters to facilitate the inference algorithm's
numerical convergence. In particular, it defines the population mean fitness
parameter $\mu_{\bar{s}_t}$ as
$$
\mu_{\bar{s}_t} = \frac{1}{N} \sum_{n=1}^N 
-\ln\left( \frac{r_{t+1}^{(n)}}{r_t^{(n)}} \right),
$${#eq-prior_meanfit_mean_data}
where $N$ is the number of neutral lineages and $r_t^{(n)}$ is the number of
neutral lineages. In other words, it defines the mean of the prior distribution
as the mean of what one naively would compute from the neutral lineages,
discarding cases where the ratio diverges because the denominator $r_t^{(n)} =
0$. For the variance parameter, we chose a value $\sigma_{\bar{s}_t} = 0.05$.

>>>>

With this in mind, we leave it to the reader to judge the selection of priors.
Furthermore, the software package associated with this work, `BarBay.jl`,
is written so that users can experiment with different prior selection criteria
that fit their needs. We strongly advocate that statistics should not be done in
a black-box fit-all tool mindset but rather as a formal way to encode the
assumptions behind the analysis, subject to constructive criticism. With this
philosophical baggage behind us, let us now focus on how the priors used for
this work were selected.

@@

The `BarBay.jl` package includes a function `naive_prior` within the
`stats` module. This function utilizes the data from the neutral lineages to
determine some of the prior parameters to facilitate the inference algorithm's
numerical convergence. In particular, it defines the population mean fitness
parameter $\mu_{\bar{s}_t}$ as
$$
\mu_{\bar{s}_t} = \frac{1}{N} \sum_{n=1}^N 
-\ln\left( \frac{r_{t+1}^{(n)}}{r_t^{(n)}} \right),
$${#eq-prior_meanfit_mean_data}
where $N$ is the number of neutral lineages and $r_t^{(n)}$ is the number of
neutral lineages. In other words, it defines the mean of the prior distribution
as the mean of what one naively would compute from the neutral lineages,
discarding cases where the ratio diverges because the denominator $r_t^{(n)} =
0$. For the variance parameter, we chose a value $\sigma_{\bar{s}_t} = 0.05$.

### 2023-10-09 doc/_bayesian_inference.qmd
change from BayesFitness to BarBay

This work is accompanied by the Julia library `BayesFitness.jl` that makes use
of the implementation of both MCMC-based integration as well as ADVI
optimization to numerically approximate the solution of @eq-bayes_full_marginal
within the Julia ecosystem [@ge2018].

>>>>

This work is accompanied by the Julia library `BarBay.jl` that makes use
of the implementation of both MCMC-based integration as well as ADVI
optimization to numerically approximate the solution of @eq-bayes_full_marginal
within the Julia ecosystem [@ge2018].

### 2023-10-09 doc/_ppc.qmd
small fixes
change from BayesFitness to BarBay

## Posterior predictive checks

@@

In practice, to compute the posterior predictive checks, we generate multiple samples from the posterior distribution $\pi(\theta \mid \underline{r}^{(m)})$ 
$$
\underline{\theta} = (\theta_1, \theta_2, \ldots, \theta_N).
$${#eq-ppc_theta_samples}
With these samples in hand, the `BayesFitness.jl` package includes the function
`logfreq_ratio_bc_ppc` for non-neutral barcodes that uses this set of posterior
parameter samples to generate samples from the distribution defined in
@eq-ppc_logfreq_likelihood. For a large-enough number of samples, we can then
compute the desired percentiles---5, 68, and 95 percentiles in all figures in 
the main text---that are equivalent to the corresponding credible regions. In
other words, the range of values of $\ln\gamma_t^{(m)}$ generated by this 
bootstrap process can be used to compute the region where we expect to find our
raw estimates $\ln\hat{\gamma}_t^{(m)}$ with the desired probability. The
package `BayesFitness.jl` includes an equivalent function,
`logfreq_ratio_popmean_ppc`, for neutral lineages.

>>>>

## Posterior predictive checks {#sec-ppc}

@@

In practice, to compute the posterior predictive checks, we generate multiple
samples from the posterior distribution $\pi(\theta \mid \underline{r}^{(m)})$ 
$$
\underline{\theta} = (\theta_1, \theta_2, \ldots, \theta_N).
$${#eq-ppc_theta_samples}
With these samples in hand, the `BarBay.jl` package includes the function
`logfreq_ratio_bc_ppc` for non-neutral barcodes that uses this set of posterior
parameter samples to generate samples from the distribution defined in
@eq-ppc_logfreq_likelihood. For a large-enough number of samples, we can then
compute the desired percentiles---5, 68, and 95 percentiles in all figures in
the main text---that are equivalent to the corresponding credible regions. In
other words, the range of values of $\ln\gamma_t^{(m)}$ generated by this
bootstrap process can be used to compute the region where we expect to find our
raw estimates $\ln\hat{\gamma}_t^{(m)}$ with the desired probability. The
package `BarBay.jl` includes an equivalent function,
`logfreq_ratio_popmean_ppc`, for neutral lineages.
